---
title: "Homework 5"
output: github_document
---

##### Import Packages

```{r}
library(ggplot2)
library(dplyr)
library(purrr)
library(tidyverse)
```

### Problem 1
This zip file contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

  - Start with a dataframe containing all file names; the list.files function will help
  - Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe
  - Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary

```{r}
# read in files in data folder
filenames <- list.files("data/problem-1", pattern="*.csv", full.names=TRUE)
filenames
#create a combines dataframe
combo_data <- map_df(filenames, read.csv,stringsAsFactors = FALSE, .id = 'filename') %>%
  mutate(filename = substr(filenames,16,21)) %>% #shorten filename variable
  mutate(condition = substr(filenames,16,18), # add condition variable
                condition = recode(condition, "con" = "control", "exp" = "experimental"),
                condition = factor(condition)) %>%
  janitor::clean_names()

#tidy the dataset, longer format
combo_tidy_data <- pivot_longer(combo_data, week_1:week_8, 
    names_to = "week", 
    names_prefix = "week_",
    values_to = "value") %>%
    mutate(week = as.numeric(week))

```

- Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.

```{r}
p <- ggplot(data = combo_tidy_data, aes(x = week, y = value, color = condition, group = filename))
p + geom_line()
```
Looking at the plot we can see differences in the trends between the control and experimental groups across weeks of the study. 

Both control and experimental particpants began the study in a similar value in week 0. Values for participants in the control group remain in a relatively consistent across the 8 weeks of the study. Values for participants in the experimental group appear to trend upward across weeks of the study.

### Problem 2

The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository.

###### Raw Data Description: 

The file consists of information for 52,179 homicide victims from 2007 to 2017. For each victim, the file contains information across 12 variables including their name, sex, the date the homicide was reported, the city, the state, region (specifically latitude and longitude), and legal case status (disposition). 

###### Preprocessing Steps: 

The data then underwent the following preprocessing steps.

- Use janitor to clean the file. Convert victim_race, 
- Create a city_state variable (e.g. “Baltimore, MD”) 
- Convert victim_race, victim_sex, and disposition into factors
- *homicides_short_df* created summarizing cities with the total number of cases, number unsolved, and proportion unsolved.

```{r}
# read in homicides file & tidy
homicides_df <- read.csv("data/problem-2/homicide-data.csv") %>%
  janitor::clean_names() %>%
  mutate(city_state= paste(city,state,sep = ", ")) %>%
  mutate(disposition = as.factor(disposition), victim_race = as.factor(victim_race),
         victim_sex = as.factor(victim_sex)) %>%

#Create df grouped by city. For each city, summarize total people, number of unsolved cases, and proportion unsolved.
select(city_state, reported_date, victim_last, victim_race, disposition) %>% 
  group_by(city_state) %>%
  summarize (
    total = n(),
    unsolved = sum(disposition == "Closed without arrest") + sum(disposition == "Open/No arrest"),
    solved = total - unsolved,
    rate = unsolved/total) -> homicides_short_df

```

Estimate the proportion of homicides unsolved in Baltimore, MD and then across all the cities in the data.

To do this, we created *prop_test_city* a function that reads in a city, uses prop.test() to estimate the proportion of unsolved homicides, uses broom::tidy to tidey the output and this object and them returns the city, estimated proportion and confidence interval.

```{r}

prop_test_city = function (element) {
  
  data = homicides_short_df$city_state==element #subset data for Baltimore, MD
  n = homicides_short_df$total[data] #extract total cases
  counts = homicides_short_df$unsolved[data] #extract unsolved cases
  
  prop.test(counts,n) %>% #calculate proportions
    broom::tidy() %>% 
    mutate(city_state = element) %>%
    select(city_state,estimate,conf.low,conf.high) -> values
  
  return(values)
}


# Use the prop.test function to estimate prop of homicides that are unsolved in Batlimore, MD
prop_test_city("Baltimore, MD")

# For cities in dataset, extract proportion of unsolved homicides and CI
homicides_CI <- purrr::map_df(homicides_short_df$city_state, prop_test_city) %>% 
  filter(city_state != "Tulsa, AL") #remove erroneous datapoint

```
Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

```{r}
#reordered dataset by estimate (descending)
homicides_CI[order(homicides_CI$estimate, decreasing = TRUE),] %>% 
  
  #not sure why resorted dataset is not getting carried forward to graph
  
ggplot(aes(x = estimate, y = city_state)) + 
  geom_point() + #add estimates
  geom_errorbar(aes(xmin=conf.low, xmax = conf.high)) +
  xlab("Unsolved Homicide Rate") +
  ylab("Location")

```

### Problem 3
When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a one-sample t-test.

First set the following design elements:

- Fix n=30
- Fix σ=5
- Set μ=0. Generate 5000 datasets from the model

x∼Normal[μ,σ]

For each dataset, save μ̂  and the p-value arising from a test of H:μ=0 using α=0.05. Hint: to obtain the estimate and p-value, use broom::tidy to clean the output of t.test.

Repeat the above for μ={1,2,3,4,5,6}, and complete the following:

1. Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ on the x axis. 

*Describe the association between effect size and power.*

2. Make a plot showing the average estimate of μ̂  on the y axis and the true value of μ on the x axis. Make a second plot (or overlay on the first) the average estimate of μ̂  only in samples for which the null was rejected on the y axis and the true value of μ on the x axis. 

*Is the sample average of μ̂  across tests for which the null is rejected approximately equal to the true value of μ? Why or why not?*
